Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-24 paper.qmd
Updated Figure 11

## Compute hexbin parameters
num_bins_x_gau1 <- 25
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < 0.1) #benchmark_gau1

distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
  mutate(ID = row_number()) |>
  dplyr::filter(!(ID %in% c(150, 110))) |>
  dplyr::select(-ID)

@@

## Set the maximum difference as the criteria
distance_df_small_edges <- distance_gau1 |>
  dplyr::filter(distance < 0.1) |>
  mutate(ID = row_number()) |>
  filter(!(ID %in% c(150, 110))) |>
  dplyr::select(-ID) #benchmark_gau1

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 50
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < 0.05) #|> benchmark_gau1
  #mutate(ID = row_number()) #|>
  #filter(!(ID %in% c(112)))

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 22
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < 0.06) #|> #benchmark_gau1
  #mutate(ID = row_number()) #|>
  #filter(!(ID %in% c(97)))

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    #alpha = 0.4,
    linewidth = 0.5) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 1.5)

>>>>

## Compute hexbin parameters
num_bins_x_gau1 <- 23
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < 0.06) #benchmark_gau1

# distance_df_small_edges_gau1 <- distance_df_small_edges_gau1 |>
#   mutate(ID = row_number()) |>
#   dplyr::filter(!(ID %in% c(150, 110))) |>
#   dplyr::select(-ID)

@@

## Set the maximum difference as the criteria
distance_df_small_edges <- distance_gau1 |>
  dplyr::filter(distance < 0.06) #|>
  #mutate(ID = row_number()) |>
  #filter(!(ID %in% c(150, 110))) |>
  #dplyr::select(-ID) #benchmark_gau1

@@

five_gau_proj_tsne_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    #size = 0.5,
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2),
    label=rownames(axes), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("a2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 75
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < benchmark_gau1) #|> benchmark_gau1
  #mutate(ID = row_number()) #|>
  #filter(!(ID %in% c(112)))

@@

five_gau_proj_umap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1, 
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df, 
    aes(
      x = proj1_from, 
      y = proj2_from, 
      xend = proj1_to, 
      yend = proj2_to), 
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes, 
    aes(x=x1, y=y1, xend=x2, yend=y2), 
    colour="grey70") +
  geom_text(
    data=axes, 
    aes(x=x2, y=y2, label=rownames(axes)), 
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle, 
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("b2", 
                      position = c(0.08, 0.9),
                      cex = 1.5)

@@

## Compute hexbin parameters
num_bins_x_gau1 <- 33
lim1 <- gau1_scaled_obj$lim1
lim2 <- gau1_scaled_obj$lim2
r2_gau1 <- diff(lim2)/diff(lim1)

@@

distance_df_small_edges_gau1 <- distance_gau1 |>
  filter(distance < benchmark_gau1) #|> #benchmark_gau1
  #mutate(ID = row_number()) #|>
  #filter(!(ID %in% c(97)))

@@

five_gau_proj_pacmap_model1 <- projected_df |>
  ggplot(
    aes(
      x = proj1,
      y = proj2)) +
  geom_point(
    alpha = 0.05,
    color = clr_choice) +
  geom_segment(
    data = model_df,
    aes(
      x = proj1_from,
      y = proj2_from,
      xend = proj1_to,
      yend = proj2_to),
    color = "#000000",
    alpha = 0.4,
    linewidth = 0.8) +
  geom_segment(
    data=axes,
    aes(x=x1, y=y1, xend=x2, yend=y2),
    colour="grey70") +
  geom_text(
    data=axes,
    aes(x=x2, y=y2, label=rownames(axes)),
    colour="grey50",
    size = 3) +
  geom_path(
    data=circle,
    aes(x=c1, y=c2), colour="grey70") +
  coord_fixed() +
  xlim(c(-0.89, 0.89)) +
  ylim(c(-0.89, 0.89)) +
  interior_annotation("c2",
                      position = c(0.08, 0.9),
                      cex = 1.5)
